A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-04-13 example_problem.tex
label fix

\section{Example Problem}
This section presents an example problem to demonstrate the process of obtaining the FPG and the advantages of doing so. The task is the conceptual sizing of a single--aisle subsonic transport using an MDAO framework with objectives for performance and total weight for a required mission. The full set of analysis codes available for use is given in Table \ref{t:analysis codes}.
\begin{table}[htbp]
  \centering
  \caption{Analysis codes for subsonic transport sizing}
    \begin{tabular}{cc}
    \toprule
    analysis code & description \\
    \midrule
    VSP   & parametric geometry \\
    PDCYL & wing and fuselage weight estimation \\
    NPSS  & engine sizing and performance \\
    VORLAX & aerodynamics using the vortex lattice method \\
    PMARC & aerodynamics using a low-order panel method \\
    WATE  & engine weight estimation \\
    FLOPSa & mission performance, \\
      & engine sizing, and weight estimation \\
    FLOPSb & mission performance only \\
    \bottomrule
    \end{tabular}%
  \label{t:analysis codes}%
\end{table}%
Each of these analysis codes contribute individual disciplinary analysis but they are not mutually exclusive. For example, VORLAX and PMARC are both aerodynamics codes that predict inviscid drag but with different levels of fidelity. The Flight Optimization System (FLOPS) is included twice to represent two different uses. FLOPSa denotes FLOPS being executed for both mission performance and engine analysis, while FLOPSb indicates that FLOPS is executed for only mission performance analysis. This sort of representation is useful for ``supercodes'' that care capable of being used in different ways and with different sets of inputs and outputs.

@@

\subsection{Obtaining an FPG}
The process of obtaining an FPG from the given MCG, but it is expected to include the user because of required changes to the MCG or make decisions

((I should show a figure of the MCG with edges and nodes colored as to when they were removed))
((mention python))

As presented in Sec.(()), the first step to obtaining an FPG is to detect holes. To start, we set the lower indegree limit for every node to be one to find any potential holes. In this case, it turns out that the variable nodes representing the number of engines is a hole for NPSS, FLOPSa, and FLOPSb. This implies than an FPG cannot be obtained from the given MCG, as shown in Fig.~\ref{f:MCG nengines}. In this figure, the green nodes indicate analysis blocks with a hole, and the cyan nodes indicate analysis blocks that became unused after the invalid analyses were removed ((connections were removed, color edges too)).
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/MCG_nengines}
  \end{center}
  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
\label{f:MCG nengines}
\end{figure}

@@

While Sec.(()) indicated the process for obtaining an FPG, it did not specify how to decide which edges to keep when resolving a collision, that freedom is left to the specific implementation. The benefit of this approach is that standard graph theory algorithms can be used to make these decisions. The first example of this is to obtain an FPG with the fewest possible number of cycles, which are found using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX. The FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, which indicates one cycle shown in yellow. A similar alternative would be to minimize the number of analysis blocks involved in cycles, counting multiplicity.
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/FPG_fewest_cycles}
  \end{center}
  \caption{FPG with the fewest number of cycles.}
\label{f:FPG fewest cycles}
\end{figure}

Alternatively, it may be desirable to resolving collisions by including the higher fidelity codes, such as NPSS in this case.
One method to ensure that the edges directed from certain analysis blocks are the ones chosen to resolve the conflicts is to use a ranking system. For this case, consider the rankings for each analysis code given in Table \ref{t:rankings}.
\begin{table}[htbp]
  \centering
  \caption{Example ranking of importance}
    \begin{tabular}{cc}
    \toprule
    analysis code & rank \\
    \midrule
    VSP   & 5 \\
    PDCYL & 5 \\
    NPSS  & 4 \\
    PMARC & 4 \\
    FLOPSb & 4 \\
    VORLAX & 3 \\
    WATE  & 3 \\
    FLOPSa & 2 \\
    \bottomrule
    \end{tabular}%
  \label{t:rankings}%
\end{table}% 
Step two in Sec.(()) (the resolution of collisions) is now conducted by selecting the edges directed from the analysis blocks with the highest rank for each collision. The resulting FPG is shown in Fig.~\ref{f:FPG highest rank}, and this graph has four cycles, which are enumerated in Table \ref{t:cycles}.
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/FPG_highest_rank}
  \end{center}
  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
\label{f:FPG highest rank}
\end{figure}
\begin{table}[htbp]
  \centering
  \caption{Cycles for the FPG obtained using ranking}
    \begin{tabular}{ccccccc}
    \toprule
    output/input & analysis code & output/input & analysis code & output/input & analysis code & output/input \\
    \midrule
    engine weight & FLOPSb & drag  & NPSS  & engine performance & WATE  & engine weight \\
    engine performance & FLOPSb & drag  & NPSS  & engine performance &       &  \\
    fuselage weight & FLOPSb & total weight & PDCYL & fuselage weight &       &  \\
    wing weight & FLOBSb & total weight & PDCYL & wing weight &       &  \\
    \bottomrule
    \end{tabular}%
  \label{t:cycles}
\end{table}%

>>>>

\section{Example Problem}
	\label{s:example problem}
	This section presents an example problem to demonstrate the process of obtaining the FPG and the advantages of doing so. The algorithm from Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} was implemented in the NetworkX package of the Python programming language. Any language with the ability to manipulate a graph should be applicable. NetworkX was chosen due to the built--in features, such as cycle detection.

	The example task is the conceptual sizing of a single--aisle subsonic transport using an MDAO framework with objectives for performance and total weight for a required mission. The full set of analysis codes available for use is given in Table \ref{t:analysis codes}.
	\begin{table}[htbp]
	  \centering
	  \caption{Analysis codes for subsonic transport sizing}
		\begin{tabular}{cc}
		\toprule
		analysis code & description \\
		\midrule
		VSP   & parametric geometry \\
		PDCYL & wing and fuselage weight estimation \\
		NPSS  & engine sizing and performance \\
		VORLAX & aerodynamics using the vortex lattice method \\
		PMARC & aerodynamics using a low-order panel method \\
		WATE  & engine weight estimation \\
		FLOPSa & mission performance, \\
		  & engine sizing, and weight estimation \\
		FLOPSb & mission performance only \\
		\bottomrule
		\end{tabular}%
	  \label{t:analysis codes}%
	\end{table}%
	Each of these analysis codes contribute individual disciplinary analysis but they are not mutually exclusive. For example, VORLAX and PMARC are both aerodynamics codes that predict inviscid drag but with different levels of fidelity. The Flight Optimization System (FLOPS) is included twice to represent two different uses. FLOPSa denotes FLOPS being executed for both mission performance and engine analysis, while FLOPSb indicates that FLOPS is executed for only mission performance analysis. This sort of representation is useful for ``supercodes'' that care capable of being used in different ways and with different sets of inputs and outputs.

@@

\subsection{Obtaining an FPG}
	\label{ss:obtaining an FPG}
	The process of obtaining an FPG from the given MCG is expected to be an iterative process in which the user applies the algorithm to determine whether or not an FPG can be obtained, and then modifies the indegree limits or the MCG to obtain a satisfactory FPG. For example, after determining the location of a hole, the user can change the indegree limit to make the node a design variable, or the user could change the MCG by adding a new analysis code such that the hole is resolved.

	As presented in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}, the first step to obtaining an FPG is to detect holes. To start, we set the lower indegree limit for every node to be unity to find any potential holes. In this case, it turns out that the variable nodes representing the number of engines is a hole for NPSS, FLOPSa, and FLOPSb. This implies than an FPG cannot be obtained from the given MCG, as shown in Fig.~\ref{f:MCG nengines}. In this figure, the green nodes indicate analysis blocks with a hole, and the cyan nodes indicate analysis blocks that became unused after the invalid analyses were removed ((connections were removed, color edges too)).
	\begin{figure}[htb!]
	  \begin{center}
		\includegraphics[width=.6\textwidth]{images/MCG_nengines}
	  \end{center}
	  \caption{The MCG with one global variable missing. The resulting holes are shown in green and unused analyses shown in cyan.}
	\label{f:MCG nengines}
	\end{figure}

@@

	While Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} indicated the process for obtaining an FPG, it did not specify how to decide which edges to keep when resolving a collision, that freedom is left to the specific implementation. The benefit of this approach is that standard graph theory algorithms can be used to make these decisions. The first example of this is to obtain an FPG with the fewest possible number of cycles, which are found using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX. The FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, which indicates one cycle shown in yellow. A similar alternative would be to minimize the number of analysis blocks involved in cycles, counting multiplicity.
	\begin{figure}[htb!]
	  \begin{center}
		\includegraphics[width=.6\textwidth]{images/FPG_fewest_cycles}
	  \end{center}
	  \caption{FPG with the fewest number of cycles.}
	\label{f:FPG fewest cycles}
	\end{figure}

	Alternatively, it may be desirable to resolving collisions by including the higher fidelity codes, such as NPSS in this case.
	One method to ensure that the edges directed from certain analysis blocks are the ones chosen to resolve the conflicts is to use a ranking system. For this case, consider the rankings for each analysis code given in Table \ref{t:rankings}.
	\begin{table}[htbp]
	  \centering
	  \caption{Example ranking of importance}
		\begin{tabular}{cc}
		\toprule
		analysis code & rank \\
		\midrule
		VSP   & 5 \\
		PDCYL & 5 \\
		NPSS  & 4 \\
		PMARC & 4 \\
		FLOPSb & 4 \\
		VORLAX & 3 \\
		WATE  & 3 \\
		FLOPSa & 2 \\
		\bottomrule
		\end{tabular}%
	  \label{t:rankings}%
	\end{table}% 
	Step two in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} (the resolution of collisions) is now conducted by selecting the edges directed from the analysis blocks with the highest rank for each collision. The resulting FPG is shown in Fig.~\ref{f:FPG highest rank}, and this graph has four cycles, which are enumerated in Table \ref{t:cycles}.
	\begin{figure}[htb!]
	  \begin{center}
		\includegraphics[width=.6\textwidth]{images/FPG_highest_rank}
	  \end{center}
	  \caption{FPG obtained using the ranking system.}
	\label{f:FPG highest rank}
	\end{figure}
	\begin{table}[htbp]
	  \centering
	  \caption{Cycles for the FPG obtained using ranking}
		\begin{tabular}{ccccccc}
		\toprule
		output/input & analysis code & output/input & analysis code & output/input & analysis code & output/input \\
		\midrule
		engine weight & FLOPSb & drag  & NPSS  & engine performance & WATE  & engine weight \\
		engine performance & FLOPSb & drag  & NPSS  & engine performance &       &  \\
		fuselage weight & FLOPSb & total weight & PDCYL & fuselage weight &       &  \\
		wing weight & FLOBSb & total weight & PDCYL & wing weight &       &  \\
		\bottomrule
		\end{tabular}%
	  \label{t:cycles}
	\end{table}%

### 2013-04-13 graph_syntax.tex

	  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
		\begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
		This represents an overload of information such that it is unclear which source to use.
	\end{description} 
	An algorithm for detecting and managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.

\section{Graph Representation of MDAO Constructs}
	\label{s:graph representation}
	%As discussed in Sec.~\ref{s:intro}, the digraph must be able to describe the individual aspects of an MDAO problem.
\subsection{Analysis Blocks and Connections}
	\label{ss:analysis blocks and connections}
	Analysis tools take in a number input variables and then perform some work to calculate 
	the values for their respective outputs. In an MDAO graph, this process is 
	represented by a group of nodes and edges called an \emph{analysis block}, 
	shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
	node represents a single input or output and is connected 
	to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
	the analysis block contains two model nodes, with a single edge connecting them. 
	This edge represents the necessary calculations to map given inputs 
	to proper output values. This edge can be given a weight to represent the computational cost of the analysis block. However, it is also allowed to skip 
	this edge and have both incomming and outgoing edges to variables through a single model 
	node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
	the edges within them are fixed, they become fixed sub-graphs within a graph.
	\begin{figure}[htb!]
		\begin{center}
		\includegraphics[width=4in]{images/analysis_block}
		\end{center}
		\vspace{-10pt}
	\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
	\label{f:analysis block}
	\end{figure}

@@

	MDAO problems require that information be passed between sets of analyses. When 
	information from the output of one analysis block is passed, or connected, to the 
	input of another analysis block, a new connection edge is added connecting the two 
	variable nodes involved in the exchange. These connection edges are free---unlike the edges 
	within an analysis block---which means they could be added or removed depending on the needs
	of a given design problem. In other words, free connection edges are created by 
	engineers linking the output of one tool to the input of another one. Note that 
	it is allowed for a single output to have outgoing connection edges to multiple 
	downstream inputs.

\subsection{Design Variables}
	\label{ss:design variables}
	((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges. It's seems possible that in the MCG a node has incoming edges but the user wants it to be a design variable. I can make the change if you agree.))
	Design variables require external decisions 
	to be made about a given system to identify them. Consider 
	a notional problem given by the graph in Fig.~\ref{f:designvars}. There are three input variables 
	which could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
	 $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
	have no incomming edges and could each be selected as design variables. 
	\begin{figure}[htb!]
	  \begin{center}
		\includegraphics[width=.6\textwidth]{images/design_vars_graph}
	  \end{center}
	  \caption{Notional graph with two potential design variables \label{f:designvars}}
	\end{figure}

	Given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, there will likely be a set of holes preventing a valid FPG from being obtained.
	Each hole represents a node that could potentially become a design variable. 
	It is up to the designer to examine each and decide if it is appropriate to 
	consider it a design variable, in which case the designer shall set the lower indegree limit to be zero. However, some variable holes may not actually be suitable as a design variable. For instance, if an aircraft mission analysis code requires a drag input 
	which ends up being a hole then it's likely best to leave
	$\txt{deg}_l^-(v)=1$ and find a new analysis code to fill the hole in graph.

	A design variable is defined is any variable node with $\txt{deg}_l^-(v)=0$. ((DP: it seems like this should be have been at the beginning of this subsection.)) 
	When working with a MCG or FPG, connection edges are the only free edges, so a design 
	variable in either graph will be a variable node ((only variable nodes that are local inputs)) with no incoming edges. For a valid PSG 
	any design variable node would have at least one, but possibly multiple, incomming 
	implicit edges from one or more driver nodes.

\subsection{Objectives, and Constraints}
	\label{ss:objectives and constraints}
	Objectives and constraints are also MDAO constructs that need to be identified by engineers 
	as part of the design process. In the case of objective functions single output values 
	could be selected, but commonly multiple values are assembeled together via simple 
	algebraic expression to form a objective function. For example, the problem depicted in 
	Fig.~\ref{f:designvars} could have an objective function of $z^2$ or $2y + z$. Constraints are 
	always given in the form of either an inequality of an equality, e.g. 
	$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer compiles an
	algebraic expression that can be evaluated to produce an output.

	These operations, though very simple, are effectively the same kind of task 
	performed by an analysis block. A constraint or objective function can therfore 
	be represented as another analysis block on the graph, with it's own input and 
	output variable nodes. Although fundamentally no different than an analysis block, 
	for clarity and convience it is useful to distinguish between regular analysis 
	blocks and those that arise from the addition of objectives or constraints to 
	the graph. Therefore, we define an \emph{expression block} as the collection of variable and model 
	nodes related to a given objective or constraint function. Fig.~\ref{f:obj-cons}
	shows a notional problem graph augmented with expression blocks and appropriate 
	connection edges. 
	\begin{figure}[htb!]
	  \begin{center}
		\includegraphics[width=.6\textwidth]{images/obj_const_graph}
	  \end{center}
	  \caption{Notional graph with and objective and constraint nodes \label{f:obj-cons}}
	\end{figure}

\subsection{Local vs Global Nodes}
	\label{ss:local vs global}
	A simple definition of a global node is any node that directly involves information 
	from more than one analysis block. Likewise, a local node is any node 
	that involves information from only a single analysis block. As stated before, variable 
	nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
	and similarly model nodes are also local. Expression and driver nodes
	are not inherently local or global though. For these, their locality is determined 
	by the incomming and outgoing edges they possess. Consider the graph in Fig. \ref{f:local_global}, 
	where the expression node $g_1$ represents some kind of constraint for the optimziation problem. 
	Since $g_1$ only has edges comming from analysis block $B$, it would be considered a local constraint 
	on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint. 
	\begin{figure}
	  \begin{center}
	  %\includegraphics[height=.25\textheight]{XDSM/simple}
	  [[insert graph here]]
	  \caption{A simple graph with local and global nodes \label{f:local_global}}
	  \end{center}
	\end{figure}

@@

\subsection{Coupling Between Analyses}
	\label{ss:coupling}
	Coupling exists in a design problem when a set of two or more analyses each depend on the
	others outputs. Consider the notional problem from Sec.~\ref{s:specific vs fundamental} where two 
	analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
	is shown in Fig.~\ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
	edge to $B.z$, indicating that there is a connection between the two. 
	Similarly a connection exists between $B.y$ and $A.y$. These two connections form 
	a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
	to be iterated on until they converge. 
	\begin{figure}
	  \begin{center}
	  %\includegraphics[height=.25\textheight]{}
	  [INSERT GRAPH HERE]
	  \caption{Graph of a notional problem with simple coupling \label{f:coupling}}
	  \end{center}
	\end{figure}

>>>>

  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
\end{description} 
These violations will be managed in order to seek a working problem formulation.

\section{Graph Representation of MDAO Constructs}
\label{s:graph representation}
In this section we will make use of the Sellar problem, described in 
Eqn. \ref{eqn:simple_fpf}. The graph representation of the Sellar problem, 
using the new stynax, is presented in Fig. \ref{f:sellar_graph_full}. Throughout
the rest of this section, different sub-graphs or slight modifications of 
the full graph will be used to demonstrate the important MDAO constructs within
this graph syntax.

\begin{figure}[htb!]
    \begin{center}
    \includegraphics[width=4in]{images/sellar_graph_full}
    \end{center}
    \vspace{-10pt}
\caption{Graph of the Sellar problem formulation}
\label{f:sellar_graph_full}
\end{figure}

%As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
\subsection{Analysis Blocks and Connections}
\label{ss:analysis blocks and connections}
Analysis tools take in a number input variables and then perform some work to calculate 
the values for their respective outputs. In an MDAO graph, this process is 
represented by a group of nodes and edges called an \emph{analysis block}, 
shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
node represents a single input or output and is connected 
to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
the analysis block contains two model nodes, with a single edge connecting them. 
This edge represents the necessary calculations to map given inputs 
to proper output values. If constructing a weighted graph, the weight calculation edge 
provides the opportunity to computational cost. However, it is also allowed to skip 
this edge and have both incomming and outgoing edges to variables through a single model 
node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
the edges within them are fixed, they become fixed sub-graphs within a graph.

\begin{figure}[htb!]
    \begin{center}
    \includegraphics[width=4in]{images/analysis_block}
    \end{center}
    \vspace{-10pt}
\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
\label{f:analysis block}
\end{figure}

@@

MDAO problems require that information be passed between sets of analyses. When 
information from the output of one analysis block is passed, or connected, to the 
input of another analysis block, a new connection edge is added connecting the two 
variable nodes involved in the exchange. These connection edges are free, unlike the edges 
within an analysis block, they could be added or removed depending on the needs
of a given design problem. In other words, free connection edges are created by 
engineers linking the output of one tool to the input of another one. Note that 
it is allowed for a single output to have outgoing connection edges to multiple 
downstream inputs.

\subsection{Design Variables}
((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges. It's seems possible that in the MCG a node has incoming edges but the user wants it to be a design variable. I can make the change if you agree.))
For the Sellar problem, there are three design variables, $x_1^*$, $z_1^*$, and $z_2^*$. In Fig. 
\ref{f:sellar_graph_full}, these are the only three holes (nodes without out 
incomming edges) in the graph. Based on this example it's tempting to simply 
call all holes in the graph design variables. However, in the general case this is 
not completely correct. For any given problem, the set of holes only identifies 
all the potential design variables. However, some variable holes might not make good design 
variables. For instance, if your aircraft mission analysis code requires a drag input 
which ends up being a hole then it's likely that some new analysis tool is needed
to predict drag before the design process can continue.

\begin{figure}[htb!]
  \begin{center}
  [NOTE: Re-work this one to match $D_2$ from Sellar]
    \includegraphics[width=.6\textwidth]{images/design_vars_graph}
  \end{center}
  \caption{Notional graph with two potential design variables \label{f:designvars}}
\end{figure}

More formally stated, given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, 
there will be a set of holes. Each hole represents a node that could potentially 
become a design variable. It is up to the designer to examine that variable and 
decide if it is appropriate to use it as a design variable. If a variable should 
be a design variable, then set it's ${deg}_l^-(v)=0$. Otherwise you leave 
${deg}_l^-(v)=1$ and find some tool to fill the hole in graph.

So a design variable is defined as any variable node with ${deg}_l^-(v)=0$. 
This, by definition, means that design variables are not holes in the graph, since 
they have not violated their lower in-degree limit. Regardless, 
when working with a MCG or FPG, only connection edges are allowed, so a design 
variable in either graph will be a variable with no incomming edges edges. For a valid PSG 
any design variable node would have at least one, but possibly multiple, incomming 
implicit edges from one or more driver nodes.

\subsection{Objectives, and Constraints}
\label{ss:objectives and constraints}
Like design variables, objectives and constraints are also constructs that need 
to be identified when setting up a given design problem. In the case of 
objective functions single output values could be selected, but commonly multiple 
values are assembeled together via simple algebraic expressions to form a composite 
objective function. (e.g. The objective function from Eqn. \ref{eqn:simple_fpf} is
$x_1^2+z_2+y_1+e^{-y_2}$). Constraints are always given in the 
form of either an inequality or an equality. (e.g. a constraint from Eqn. 
\ref{eqn:simple_fpf} is $\frac{y_1}{3.16}-1\geq0$.) By convention, 
constraints are usually given such that some expression will be less than or equal 
to zero, so any positive value would violate the constraint. For both objective 
and constraints, these simple expressions take some inputs and map them to an
output value of significance to the overall design problem.

These operations, though very simple, are effectively the same kind of task 
performed by an analysis block. A constraint or objective function can therfore 
be represented as another analysis block on the graph, with it's own input and 
output variable nodes. Although fundamentally no different than an analysis block, 
for clarity and convience it is useful to distinguish between regular analysis 
blocks and those that arise from the addition of objectives or constraints to 
the graph. So we define an expression block, as the collection of variable and model 
nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
shows a notional problem graph augmented with expression blocks and appropriate 
connection edges.

\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/obj_const_graph}
  \end{center}
  \caption{Notional graph with and objective and constraint nodes \label{f:obj-cons}}
\end{figure}

\subsection{Local vs Global Nodes}

  A simple definition of a global node is any node that involves information 
  from more than one analysis block. Likewise, a local node is any node 
  that involves information from only a single analysis block. As stated before, variable 
  nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
  and similarly model nodes are also local. Expression and driver nodes
  are not inherently local or global though. For these, their locality is determined 
  by the incomming and outgoing edges they possess. Consider the graph in Fig. \ref{f:local_global}, 
  where the expression node $g_1$ represents some kind of constraint for the optimziation problem. 
  Since $g_1$ only has edges comming from analysis block $B$, it would be considered a local constraint 
  on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint.

  \begin{figure}
      \begin{center}
      %\includegraphics[height=.25\textheight]{XDSM/simple}
      [[insert graph here]]
      \caption{A simple graph with local and global nodes \label{f:local_global}}
      \end{center}
  \end{figure}

@@

\subsection{Coupling Between Analyses}

  Coupling exists in a design problem when a set of two or more analyses each depend on the
  others outputs. Consider the notional problem from section \ref{sec:spg_vs_fpg} where two 
  analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
  is shown in Fig. \ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
  edge to $B.z$, indicating that there is a connection between the two. 
  Similarly a connection exists between $B.y$ and $A.y$. These two connections form 
  a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
  to be iterated on untill they converge.

  \begin{figure}
      \begin{center}
      %\includegraphics[height=.25\textheight]{}
      [INSERT GRAPH HERE]
      \caption{Graph of a notional problem with simple coupling \label{f:coupling}}
      \end{center}
  \end{figure}

### 2013-04-13 intro.tex

\section{Specific vs Fundamental Problem Formulation }
	\label{s:specific vs fundamental}
    The Fundamental Problem Formulation (FPF) for any given problem will be constant regardless 
    of which MDAO framework, optimization algorithm, iterative solver, or solution strategy
    is used to solve the problem. For example, consider the following notional problem:

    \begin{align}
        given & \ \ A: {x,y} \rightarrow {m,z} \notag
        \\      & \ \ B: {x,z} \rightarrow {y^t} \notag
        \\min. &\ \ f(m) \notag
        \\w.r.t. & \ \ x,y,z \notag
        \\s.t. & \ \ g(y^t,y) = 0
        \label{eqn:simple}
    \end{align}

    $A$ and $B$ represent analysis tools, and $f$ and $g$ are the objective and constraint functions respectively. 
    Equation \ref{eqn:simple} makes an inherent assumption about the solution strategy for the problem. 
    Analysis $A$ outputs $z$, which is an input to $B$. Hence, $A$ should be run before $B$ with 
    with $y$ being iterated on to convergence with $y^t$. However, a slightly different formulation is 
    equally valid and still represents the exact same problem:

    \begin{align}
        given & \ \ B: {x,z} \rightarrow {y} \notag
        \\      & \ \ A: {x,y} \rightarrow {m,z^t} \notag
        \\min. &\ \ f(m) \notag
        \\w.r.t. & \ \ x,y,z \notag
        \\s.t. & \ \ g(z^t,z) = 0
        \label{eqn:simple2}
    \end{align}

    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
    and $z$ will be iterated on to convergence with $z^t$. Now the problem can be solved by running $B$ first and then $A$.
    Since the formulations in Eqs. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem and neither can be the
    FPF. They are both specific versions of the same more fundamental description of 
    the problem that is common between them. We present the FPF as follows:

    \begin{align}
        given & \ \ A: {x,y} \rightarrow {m,z^t} \notag
        \\      & \ \ B: {x,z} \rightarrow {y^t} \notag
        \\min. &\ \ f(m) \notag
        \\w.r.t. & \ \ x,y,z \notag
        \\s.t. & \ \ g1(z^t,z) = 0 \notag
        \\     & \ \ g2(y^t,y) = 0
        \label{eqn:simple_fpf}
    \end{align}

    The FPF in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} and Eq. \ref{eqn:simple2} because it has 
    two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
    either $A$ or $B$ could be run first or both could be run simultaneously. Alexandrov and Lewis investigated problem 
    formulation in this fully decoupled way as part of their work on 
    Reconfigurability in MDO Problem Synthesis (REMS) \cite{alexandrov2004}.
    In this work they show how removing either constraint and replacing 
    it with a direct dependence between the two analyses, a process they term ``collapsing'', 
    you could regain the earlier two problem formulations.

>>>>

\section{Specific vs Fundamental Problem Formulation }
	\label{s:specific vs fundamental}
    The Fundamental Problem Formulation (FPF) for any given problem will be constant regardless 
    of which MDAO framework, optimization algorithm, iterative solver, or solution strategy
    is used to solve the problem. However, it is possible to formulate a problem
    with an assumed execution ordering that implies a solution strategy. 
    For example, consider the Sellar test problem \cite{AIAA:sellar}:

    \begin{align}
        given & \ \ y_1 = D_1(x_1,y_2,z_1,z_2) \notag
        \\      & \ \ y_2^t = D_2(y_1,z_1,z_2) \notag
        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
        \\s.t. & \ \ G(y_2,y_2^t) = 0
        \label{eqn:simple}
    \end{align}

    $D_1$ and $D_2$ represent analysis tools, and $F$ and $G$ are the objective 
    and constraint functions respectively. Equation \ref{eqn:simple} makes an 
    inherent assumption about the solution strategy for the problem. Analysis 
    $D_1$ outputs $y_1$, which is an input to $D_2$. Hence, $D_1$ should be run 
    before $D_2$ with with $y_2$ being iterated on to convergence with $y_2^t$. 
    However, a slightly different formulation is equally valid and still 
    represents the exact same problem:

    \begin{align}
        given & \ \ y_2 = D_2(y_1,z_1,z_2) \notag
        \\      & \ \ y_1^t = D_1(x_1,y_2,z_1,z_2) \notag
        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
        \\s.t. & \ \ G(y_1,y_1^t) = 0
        \label{eqn:simple}
    \end{align}

    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. 
    $D_1$ is now dependent on the output of $D_2$, and $y_1$ will be iterated on to 
    convergence with $y_1^t$. Now the problem can be solved by running $D_2$ first 
    and then $D_1$. Since the formulations in Eqs. \ref{eqn:simple} and 
    \ref{eqn:simple2} both describe the same problem then neither can be the
    FPF. They are both specific versions of a more fundamental description of 
    the problem that is common between them. We present the fundamental 
    formulation as follows:

    \begin{align}
        given & \ \ D_1.y_1 = D_1(x_1,y_2,z_1,z_2) \notag
        \\      & \ \ D_2.y_2 = D_2(y_1,z_1,z_2) \notag
        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
        \\s.t. & \ \ G_1(D_1.y_1,D_2.y_1) = 0 \notag
        \\     & \ \ G_2(D_1.y_2,D_2.y_2) = 0
        \label{eqn:simple_fpf}
    \end{align}

    The formulation in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} 
    and Eq. \ref{eqn:simple2} because it has two constraints which both must be 
    met. The presence of both of these constraints fully decouples the problem so that 
    either $D_1$ or $D_2$ could be run first or both could be run simultaneously. 
    Alexandrov and Lewis, in their work with REMS, assert that fully decoupling 
    the analyses is key to achieving reconfigurability in the problem 
    formulation\cite{alexandrov2004}. The key feature they identify to allow this decoupling 
    is the creation of separate variables for each analysis so that a variable 
    is identified by its analysis and it's name. For example, in Eqn. \ref{eqn:simple}
    there is only one $y_1$ variable vs Eqn. \ref{eqn:simple_fpf} where 
    there are two ($D_1.y_1$ vs $D_2.y_1$).
